### 2024-06-14 paper/paper.tex
Initial commit

\PassOptionsToPackage{table}{xcolor}
\documentclass[review,onefignum,onetabnum,a4paper]{siamart190516}
\pdfoutput=1
\overfullrule=0pt

%------------------------------------------------------------------------------
% Header
%------------------------------------------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath, amssymb, gensymb, physics, upgreek, siunitx, bbm}
\usepackage{tikz, pgfplots, graphicx, epstopdf, float, hyperref, xspace}
\usepackage[caption=false]{subfig}
\usepackage{enumitem, verbatim, listings}
\usepackage{mathtools, stmaryrd}
\usepackage{booktabs, array, ragged2e}
\usepackage[export]{adjustbox}
\usepackage{cleveref}

\sisetup{table-number-alignment=center, exponent-product=\times}
\newcommand{\crefrangeconjunction}{--}

\usetikzlibrary{trees, cd, babel}
\graphicspath{{./figures}}
\epstopdfsetup{outdir=./}

\lstset{basicstyle=\footnotesize\ttfamily}

\definecolor{paired1}{HTML}{a6cee3}
\definecolor{paired2}{HTML}{1f78b4}
\definecolor{paired3}{HTML}{b2df8a}
\definecolor{paired4}{HTML}{33a02c}
\definecolor{paired5}{HTML}{fb9a99}
\definecolor{paired6}{HTML}{e31a1c}
\definecolor{paired7}{HTML}{fdbf6f}
\definecolor{paired8}{HTML}{ff7f00}
\definecolor{paired9}{HTML}{cab2d6}
\definecolor{paired10}{HTML}{6a3d9a}

\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\pgfplotsset{
  log x ticks with fixed point/.style={
      xticklabel={
        \pgfkeys{/pgf/fpu=true}
        \pgfmathparse{exp(\tick)}%
        \pgfmathprintnumber[fixed relative, precision=3]{\pgfmathresult}
        \pgfkeys{/pgf/fpu=false}
      }
  },
  log y ticks with fixed point/.style={
      yticklabel={
        \pgfkeys{/pgf/fpu=true}
        \pgfmathparse{exp(\tick)}%
        \pgfmathprintnumber[fixed relative, precision=3]{\pgfmathresult}
        \pgfkeys{/pgf/fpu=false}
      }
  }
}

% Classic symbols
\newcommand{\reals}{\mathbb{R}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\dx}{\,\d\mathbf{x}}
\newcommand{\md}{\,\d}
\newcommand{\bigo}[1]{\mathcal{O}(#1)}
\DeclareMathOperator*{\argmin}{arg\,min}

\let\grad\undefined
\let\curl\undefined
\let\div\undefined
\let\nullop\undefined
\let\d\undefined
\let\tr\undefined
\DeclareMathOperator{\grad}{grad}
\DeclareMathOperator{\curl}{curl}
\DeclareMathOperator{\div}{div}
\DeclareMathOperator{\nullop}{null}
\DeclareMathOperator{\d}{d}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\Exists}{\exists}
\DeclareMathOperator{\Forall}{\forall}

\DeclareMathOperator{\refgrad}{\widehat{\grad}}
\DeclareMathOperator{\refcurl}{\widehat{\curl}}
\DeclareMathOperator{\refdiv}{\widehat{\div}}

\newcommand{\Hgrad}{H(\grad)}
\newcommand{\Hcurl}{H(\curl)}
\newcommand{\Hdiv}{H(\div)}
\newcommand{\Ltwo}{L^2}

\newcommand{\mesh}{\mathcal{T}_h}
\newcommand{\skeleton}{\mathcal{E}_h}
\newcommand{\refline}{\hat{\mathcal{I}}}
\newcommand{\jac}{\mathrm{D}F}
\newcommand{\pullback}{\mathcal{F}}
\newcommand{\betah}{\tilde{\beta}}

% Vector notation
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\uvec}[1]{\mathbf{\hat{#1}}}
\newcommand{\conj}[1]{{\overline{#1}}}
\DeclareMathOperator{\blockdiag}{blockdiag}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\spn}{span}
\DeclareMathOperator{\vecop}{vec}
\newcommand{\eye}{{\mathbb{I}}}
\newcommand{\kron}{\otimes}
\newcommand{\jump}[1]{\left\llbracket #1 \right\rrbracket}
\newcommand{\avg}[1]{\left\{ #1 \right\}}
\newcommand{\broken}[1]{\tilde{#1}}

\newcommand{\be}{\vec{e}}
\newcommand{\bu}{\vec{u}}
\newcommand{\bn}{\vec{n}}
\newcommand{\bm}{\vec{m}}
\newcommand{\bv}{\vec{v}}
\renewcommand{\bf}{\vec{f}}
\newcommand{\bq}{\vec{q}}
\newcommand{\br}{\vec{r}}
\newcommand{\bx}{\vec{x}}
\newcommand{\bw}{\vec{w}}
\newcommand{\by}{\vec{y}}
\newcommand{\bz}{\vec{z}}
\newcommand{\bB}{\vec{B}}
\newcommand{\bC}{\vec{C}}
\newcommand{\bE}{\vec{E}}
\newcommand{\bF}{\vec{F}}
\newcommand{\bJ}{\vec{J}}
\newcommand{\bP}{\vec{P}}
\newcommand{\bT}{\vec{T}}
\newcommand{\bX}{\vec{X}}

\newcommand{\ub}{\underline{b}}
\newcommand{\uc}{\underline{c}}
\newcommand{\ue}{\underline{e}}
\newcommand{\uf}{\underline{f}}
\newcommand{\ug}{\underline{g}}
\newcommand{\up}{\underline{p}}
\newcommand{\uq}{\underline{q}}
\newcommand{\ur}{\underline{r}}
\newcommand{\uu}{\underline{u}}
\newcommand{\uv}{\underline{v}}
\newcommand{\uw}{\underline{w}}
\newcommand{\ux}{\underline{x}}
\newcommand{\uy}{\underline{y}}
\newcommand{\uz}{\underline{z}}
\newcommand{\ut}{\underline{t}}
\newcommand{\ulambda}{\underline{\lambda}}

\newcommand{\uB}{\underline{B}}
\newcommand{\ubb}{\underline{\vec{b}}}
\newcommand{\ubq}{\underline{\vec{q}}}
\newcommand{\ubu}{\underline{\vec{u}}}
\newcommand{\ubv}{\underline{\vec{v}}}
\newcommand{\ubf}{\underline{\vec{f}}}
\newcommand{\ubg}{\underline{\vec{g}}}

\newcommand{\xhat}{\hat{x}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\shat}{\hat{s}}
\newcommand{\Khat}{\hat{K}}
\newcommand{\Bhat}{\hat{B}}
\newcommand{\Ahat}{\hat{A}}
\newcommand{\Dhat}{\hat{D}}
\newcommand{\Shat}{\hat{S}}
\newcommand{\Ghat}{\hat{G}}

\renewcommand{\P}{\mathrm{P}}
\newcommand{\DP}{\mathrm{DP}}
\newcommand{\Q}{\mathrm{Q}}
\newcommand{\DQ}{\mathrm{DQ}}
\newcommand{\RTCF}{\mathrm{RTCF}}
\newcommand{\RTCE}{\mathrm{RTCE}}
\newcommand{\NCF}{\mathrm{NCF}}
\newcommand{\NCE}{\mathrm{NCE}}

%\newtheorem{theorem}{Theorem}
\newtheorem{remark}{Remark}

%------------------------------------------------------------------------------
% Document
%------------------------------------------------------------------------------

\title{
   Fast solvers for the high-order FEM simplicial de Rham complex
\thanks{Submitted to the editors XX/YY/2024.
      \funding{
         PDB and PEF were supported by EPSRC grant EP/W026260/1.
      }
   }
}

\author{Pablo D.\ Brubeck\thanks{
Mathematical Institute,
University of Oxford,
Oxford, UK (\email{brubeckmarti@maths.ox.ac.uk})
%\orcid{0000-0002-3824-0080}
}
\and Patrick E.\ Farrell\thanks{
Mathematical Institute,
University of Oxford,
Oxford UK (\email{patrick.farrell@maths.ox.ac.uk})
%\orcid{0000-0002-1241-7060}
}
}

\headers{Patch solvers for simplicial $p$-FEM}{P.~D.~Brubeck and P.~E.~Farrell and R.~C.~Kirby}

\begin{document}

\numberwithin{equation}{section}
\maketitle

\begin{abstract}
This is the abstract.
\end{abstract}

\begin{keywords}
   preconditioning, de Rham complex, simplex, high-order, additive Schwarz
\end{keywords}

\begin{AMS}
   65F08, 65N35, 65N55
\end{AMS}

%\begin{DOI}
%10.1137/21XXXXXXX
%\end{DOI}

\section{Introduction} \label{sec:introduction}

\section{Conclusion} \label{sec:conclusion}

\bibliographystyle{siamplain}
\bibliography{references}
\end{document}